---
title: "Advanced R Final Project 2020"
author: "Vivekanand Kulkarni - (19200231)"
date: "15 August 2020"
output:
  html_document: default
  pdf_document: default
editor_options: 
  chunk_output_type: inline
runtime: shiny
---

# Questions:

1. Import the dataset exo_data.csv as a tibble.  Columns 2, 14 should be factors. Column 15 should be integers. The
remaining columns should be doubles.
```{r cache=FALSE, results=FALSE, warning=FALSE, comment=FALSE, message=FALSE}
# Importing required Libraries and reading csv
library(readr)
library(dplyr)
library(magrittr)
library(ggplot2)
library(shiny)
library(rstan)
library(shinystan)
library(ggiraph)
library(gganimate)
library(reshape2)
library(tidyr)
library(ggmap)
Exo_data <- read_csv("exo_data.csv")
```
Columns 1, 16, 17, 18, 25 should be
characters.Columns 2, 14 should be factors. 
Column 15 should be integers. 
The remaining columns should be doubles.
```{r cache=FALSE, results=FALSE, warning=FALSE, comment=FALSE, message=FALSE}
# Checking the structure of the data
str(Exo_data)
```
```{r cache=FALSE, results=FALSE, warning=FALSE, comment=FALSE, message=FALSE}
# Columns 1 is a character.
typeof(Exo_data$id)=="character"

#Column 16 is a character
typeof(Exo_data$recency)=="character"

#Column 17 is a character
typeof(Exo_data$r_asc)=="character"

#Column 18 is a character
typeof(Exo_data$decl)=="character"

#Column 25 is a character
typeof(Exo_data$lists)=="character"

#Column 2= flag is made a factor
Exo_data$flag %<>% as.factor

#Column 14 is made a factor
Exo_data$meth %<>% as.factor

#Column 15= year is made integer
Exo_data$year %<>% as.integer

```
2. Exclude the exoplanets with an unknown method of discovery.
```{r cache=FALSE, results=FALSE, warning=FALSE, comment=FALSE, message=FALSE}
Exo_data <- drop_na(Exo_data, meth)
```

3. Create a graphic which illustrates the relationship between the log-distances from
the Sun and the methods of discovery.
```{r cache=FALSE, results=FALSE, warning=FALSE, comment=FALSE, message=FALSE}
ggplot(Exo_data, aes(x = log(dist)))+ theme_light()+ geom_histogram(aes(color = meth, fill= meth), position = "identity", bins = 50, alpha=0.5)+ labs(title ="Log-Dist from Sun Graphic",x="Log-Dist from Sun (parsec)", y = "Count")
```
4. Create scatterplots of the log-mass versus log-distances, separating by methods of
discovery. Hovering with the cursor highlights the point and displays its name, and,
if you click, the exoplanet’s page on the Open Exoplanet Catalogue will be opened.
(paste the id after http://www.openexoplanetcatalogue.com/planet/ ).
```{r cache=FALSE, results=FALSE, warning=FALSE, comment=FALSE, message=FALSE}
Exo_data$onclick <- sprintf("window.open(\"%s%s\")",
                        "http://www.openexoplanetcatalogue.com/planet/",
                        Exo_data$id)

gg_graph = ggplot(Exo_data,
                  aes(x = log(mass),
                      y = log(dist),
                      color = meth))+theme_dark()+
                  xlab('Log(Mass)') +
                  ylab('Log(Dist') +
                  scale_color_discrete(name="Discovery Methods")+
                  geom_point_interactive(aes(data_id = id,
                             tooltip = id,
                             onclick = onclick))+
                  labs(title="Scatterplots of Log-mass vs. Log-distances")

ggiraph(code = print(gg_graph))
```
5. Rename the radius into jupiter_radius, and create a new column called earth_radius
which is 11.2 times the Jupiter radius.
```{r cache=FALSE, results=FALSE, warning=FALSE, comment=FALSE, message=FALSE}
# rename() function from tidyverse with radius into jupiter_radius.
Exo_data <- Exo_data %>% 
          rename(jupiter_radius = radius ) 
Exo_data <- Exo_data %>%
          mutate(earth_radius = jupiter_radius * 11.2 )
```

6. Focus only on the rows where log-earth radius and log-period have no missing values,
and perform kmeans with four clusters on these two columns.
```{r cache=FALSE, results=FALSE, warning=FALSE, comment=FALSE, message=FALSE}
data_clustering <- Exo_data # create new df for clustering from data

# Focus only on the rows where radius of Earth and period have no missing values
data_clustering <- Exo_data  %>% drop_na(earth_radius, period)  # 2732 obs

#log-radius of Earth and log-period
data_clustering <- data_clustering %>%
                      mutate(LogERadius = log(earth_radius),
                             LogPeriod  = log(period))


# data to perform Kmeans
data_kmeans <- data_clustering %>%
                      select(LogERadius,LogPeriod)

# perform k-means
set.seed(123)
cluster_kmeans  <- kmeans(data_kmeans, 4)
```

7.Add the clustering labels to the dataset through a new factor column called type,
with levels rocky, hot_jupiters, cold_gas_giants, others; similarly to https:
//en.wikipedia.org/wiki/Exoplanet#/media/File:ExoplanetPopulations-20170616.
png and produce the scatterplot highlighting these clusters.
```{r cache=FALSE, results=FALSE, warning=FALSE, comment=FALSE, message=FALSE}
cluster_kmeans$cluster <- as.factor(cluster_kmeans$cluster)

ggplot(data_kmeans, aes(LogPeriod,LogERadius ,color = cluster_kmeans$cluster)) + geom_point() +
                                      labs(title="Clustering solutions of Exoplanets")

# Using https://en.wikipedia.org/wiki/Exoplanet#/media/File:ExoplanetPopulations-20170616.png we have:
# 1 = cold_gas_giants 
# 2 = others 
# 3 = Rocky 
# 4 = hot_jupiters 

data_clustering$type <- cluster_kmeans$cluster
data_clustering$type <- as.numeric(data_clustering$type)

data_clustering$type[data_clustering$type == 1] <- "cold_gas_giants"
data_clustering$type[data_clustering$type == 2] <- "others"
data_clustering$type[data_clustering$type == 3] <- "Rocky"
data_clustering$type[data_clustering$type == 4] <- "hot_jupiters"

table(cluster_kmeans$cluster)
## checking
table(data_clustering$type)
```

8.Use a violin plot to illustrate how these clusters relate to the log-mass of the exoplanet.
```{r cache=FALSE, results=FALSE, warning=FALSE, comment=FALSE, message=FALSE}
# Histogram
ggplot(data_clustering, aes(x = log(mass))) +
                            geom_histogram(aes(color = type, fill = type), 
                                            position = "identity", bins = 30, alpha = 0.4) +
                                                       labs(title="Histogram of the log-mass of the exoplanet")

# Violin plot
ggplot(data_clustering, aes(x = type, y = log(mass))) + 
  geom_violin(aes(fill = type)) + labs(title="Violin plot of log-mass of the exoplanet")
```

9. Transform r_asc and decl into two new variables that are the same varibales but
in values of seconds. Use these as coordinates to represent a celestial map for the
exoplanets.
```{r cache=FALSE, results=FALSE, warning=FALSE, comment=FALSE, message=FALSE}
head(data_clustering$r_asc)
head(data_clustering$decl)
library(lubridate)
data_clustering$r_asc <- gsub(" ", ":", data_clustering$r_asc, fixed=TRUE) # convert to hh:mm:ss
data_clustering$r_asc <- hms(data_clustering$r_asc)
data_clustering$r_asc_sec <- period_to_seconds(data_clustering$r_asc)

# convert Declination to seconds and save as decl_sec
data_clustering$decl <- gsub(" ", ":", data_clustering$decl, fixed=TRUE) 
# convert to dd:mm:ss, where dd=3600ss
data_clustering$decl <- hms(data_clustering$decl) 
# for Decl, dd is similar to hh where :=3600ss
data_clustering$decl_sec <- period_to_seconds(data_clustering$decl)
# scatter plot represents a celestial map for the exoplanets
p<- ggplot(data_clustering, aes(r_asc_sec, decl_sec, color= meth)) +
                            geom_point() + 
                        labs(title="Celestial Map for Exoplanets", x="Right ascension (seconds)",
                             y="declination (seconds)")
p+xlim(c(0, 80000))
```
10. Create an animated time series where multiple lines illustrate the evolution over
time of the total number of exoplanets discovered for each method up to that year.
```{r}
ts.data <- Exo_data %>% group_by(meth, year) %>%  summarise(Count = length(meth)) %>%
                          mutate(Count = cumsum(Count))
ts.data <- na.omit(ts.data)
ggplot(ts.data, aes(x = year, y = Count, group = meth)) + 
  geom_line(aes(color = meth)) + 
  geom_point(size = 2) + 
  transition_reveal(year) + 
  labs(title = 'Evolution Total number of exoplanets discovered by methods', y = 'Number Discovered') 
 
```
11. Create an interactive plot with Shiny where you can select the year (slider widget,
with values 2009) and exoplanet type. Exoplanets appear as points on a scatterplot
(log-mass vs log-distance coloured by method) only if they have already been
discovered. If type is equal to all all types are plotted together.
```{r cache=FALSE, results=FALSE, warning=FALSE, comment=FALSE, message=FALSE}
library(shiny)

# Define UI for application that draws a histogram
ui <- fluidPage(
    
        
            sliderInput("YearSlider",
                        "Select Year", min=2009,
                        max(data_clustering$year,na.rm = TRUE),max(data_clustering$year,na.rm = TRUE),
                        round = TRUE, sep = '', width = '100%',step = 1,ticks = FALSE)
        ,
        selectInput(inputId = "exo_type",
                    label = "Type",
                    choices = c('hot_jupiters','cold_gas_giants','Rocky','others',"all"),
                    selected = 0),
        
       
      
        

        # Show a plot of the generated distribution
        
plotOutput("scatter1",height = 450)
        
    
)

# Define server logic required to draw a histogram
server <- function(input, output) {
    
  #organize data, only include planets with known mass
  
    
  
    #create scatterplot of all year's discoveries thru selected years
    output$scatter1 = renderPlot({
      if(input$exo_type=="all")
      {
        scatter1data<-data_clustering %>%filter(., year <= input$YearSlider, !is.na(mass),
                                     !is.na(dist))
      }
      else
      {
        scatter1data<-data_clustering %>%filter(., year <= input$YearSlider, !is.na(mass),
                                     !is.na(dist)) %>% filter(., type == input$exo_type)
      }
      
        
        ggplot(scatter1data, aes(x = log(dist), y = log(mass),
                                 color = meth)) +
            geom_point(size = 3) +
            labs(title = paste("Discoveries Through", input$YearSlider),
                 x = "Distance of Exoplanet Star from Sun (Parsecs)",
                 y = "Planet Mass (Jupiter Masses)",
                 colour = NULL) 
            
    })
  
}

# Run the application 
shinyApp(ui = ui, server = server)
```
12. Fit a linear regression model where log period is the response variable and the logs
of host_mass, host_temp and axis are the covariates (exclude rows that contain
at least one missing value). Include an intercept term in the regression model
```{r  warning=FALSE}
regdata <- Exo_data[,c("period","host_mass","host_temp","axis")] 
#Data with missing values are omitted
regdata.complete <- na.omit(regdata)
#We fit the linear model 
fit<-lm(log(period)~log(host_mass)+log(host_temp)+log(axis),data = regdata)
summary(fit)

```
Interpretation:


14. Embed the Shiny app from (11) in your RMarkdown document.
```{r , echo=FALSE}
library(shiny)

# Define UI for application that draws a histogram
ui <- fluidPage(
    
        
            sliderInput("YearSlider",
                        "Select Year", min=2009,
                        max(data_clustering$year,na.rm = TRUE),max(data_clustering$year,na.rm = TRUE),
                        round = TRUE, sep = '', width = '100%',step = 1,ticks = FALSE)
        ,
        selectInput(inputId = "exo_type",
                    label = "Type",
                    choices = c('hot_jupiters','cold_gas_giants','Rocky','others',"all"),
                    selected = 0),
        
       
      
        

        # Show a plot of the generated distribution
        
plotOutput("scatter1",height = 450)
        
    
)

# Define server logic required to draw a histogram
server <- function(input, output) {
    
  #organize data, only include planets with known mass
  
    
  
    #create scatterplot of all year's discoveries through the selected years
    output$scatter1 = renderPlot({
      if(input$exo_type=="all")
      {
        scatter1data<-data_clustering %>%filter(., year <= input$YearSlider, !is.na(mass),
                                     !is.na(dist))
      }
      else
      {
        scatter1data<-data_clustering %>%filter(., year <= input$YearSlider, !is.na(mass),
                                     !is.na(dist)) %>% filter(., type == input$exo_type)
      }
      
        
        ggplot(scatter1data, aes(x = log(dist), y = log(mass),
                                 color = meth)) +
            geom_point(size = 3) +
            labs(title = paste("Discoveries Through", input$YearSlider),
                 x = "Distance of Exoplanet Star from Sun (Parsecs)",
                 y = "Planet Mass (Jupiter Masses)",
                 colour = NULL) 
            
    })
  
}

# Run the application 
shinyApp(ui = ui, server = server)
```

 
